### 2021-09-29 papers/modeling-of-combined-lfr-csp-sco2-paper/backmatter.tex
tidy up figures, front and backmatter

\authorcontributions{Conceptualization, B.L., M.W., C.S. and T.N.; methodology, B.L., M.W., C.S., T.N. and B.L.; software, B.W.; validation, B.L and M.W.; formal analysis, B.W.; investigation, B.W.; resources, M.W. and B.L.; data curation, B.W.; writing---original draft preparation, B.W.; writing---review and editing, B.L. and M.W.; visualization, B.W.; supervision, M.W and B.L.; project administration, B.L.; funding acquisition, B.L, M.W, C.S and T.N. All authors have read and agreed to the published version of the manuscript.}

\funding{This work was supported by funding received from the DOE Office of Nuclear Energy’s Nuclear Energy University Program under contract number DE-NE0008988. }

% \institutionalreview{In this section, please add the Institutional Review Board Statement and approval number for studies involving humans or animals. Please note that the Editorial Office might ask you for further information. Please add ``The study was conducted according to the guidelines of the Declaration of Helsinki, and approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).'' OR ``Ethical review and approval were waived for this study, due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans or animals. You might also choose to exclude this statement if the study did not involve humans or animals.}

% \informedconsent{Any research article describing a study involving humans should contain this statement. Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans. You might also choose to exclude this statement if the study did not involve humans.

% Written informed consent for publication must be obtained from participating patients who can be identified (including by the patients themselves). Please state ``Written informed consent has been obtained from the patient(s) to publish this paper'' if applicable.}

\dataavailability{Supporting data for this study is available on request.}

\acknowledgments{We would like to thank Josh McTigue (NREL) for useful technical discussions on the development of the cycle configurations.}

\conflictsofinterest{The authors declare no conflict of interest. The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, or in the decision to publish the results'.}

%% Optional
% \sampleavailability{Samples of the compounds ... are available from the authors.}